---
title: "Logistic regression"
author: "Dat Quoc Vuong"
date: "4/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Multinomial logistic regression

```{r}
# Helper packages
library(dplyr)     # for data wrangling
library(ggplot2)   # for awesome plotting
library(rsample)   # for data splitting

# Modeling packages
library(caret)     # for logistic regression modeling

# Model interpretability packages
library(vip)       # variable importance
```

Example: Entering high school students make program choices among general program, vocational program and academic program. Their choice might be modeled using their writing score and their social economic status.

# Description of the data

```{r}
ml <- read.dta("https://stats.idre.ucla.edu/stat/data/hsbdemo.dta")
```

The data set contains variables on 200 students. 
The outcome variable is **prog**, program type. 
The predictor variables are social economic status, **ses**, a three-level categorical variable and writing score, **write**, a continuous variable. 

Let’s start with getting some descriptive statistics of the variables of interest.

```{r}
with(ml, table(ses, prog))

with(ml, do.call(rbind, tapply(write, prog, function(x) c(M = mean(x), SD = sd(x)))))
```
## Split data

```{r}
df <- attrition %>% mutate_if(is.ordered, factor, ordered = FALSE)

# Create training (70%) and test (30%) sets for the 
# rsample::attrition data.
set.seed(123)  # for reproducibility
churn_split <- initial_split(df, prop = .7, strata = "Attrition")
churn_train <- training(churn_split)
churn_test  <- testing(churn_split)
```






